Kindly Bent to Free Us

### 2019-07-08 kindly/infer/infer.tex
State the theorems.

\TODO{State the theorems}

>>>>

This section only states the theorems (and the necessary definitions).
Details and proofs are provided in \cref{appendix:constraints,appendix:infer}.

\subsubsection{Constraint system}

We requires three properties from our constraint solving algorithms: it should
compute normal firms, such normal forms should be unique, and the
system should be regular.

\begin{property}[Principal normal form]
  Normalization computes principal normal forms for $\CL$, i.e. 
  given a constraint $D\in\CL$, a substitution $\phi$ and
  $(C,\unif) = \normalize{D}{\phi}$,
  then $\phi\leq\unif$,
  $C \equivC \unif D$ and
  $\unif C = C$.
\end{property}

\begin{property}[Principal constraints]
  $\CL$ has the principal constraint property, i.e.
  for every $D\in\CL$, and a substitution $\unif$,
  either $D$ does not have a normal form, or it has
  a principal normal form.
\end{property}

\begin{property}[Regular constraint system]
  $\CL$ is regular, ie, for $x, x'$ two types or kinds,
  $\entail{}{\Ceq{x}{x'}}$ implies
  $\fv{x} = \fv{x'}$
\end{property}

The proofs are given in \cref{appendix:constraints}. These three properties
are sufficient to state that $HM(\CL)$ provides principal type inference.
However, we do not use HM(X) directly, but an extension with kind inference
and affine types and borrows.
We now show that HM(X)'s completeness and soundness theorems are still valid on
this extended system.

\subsection{Inference algorithm}

The first theorem states that our inference algorithm is sound
with respect to the syntax directed type system.

\begin{theorem}[Soundness of inference]
  Given a type environment $\E$ and a term $e$,\\
  if $\inferW{\Sv}{(C,\unif)}{\E}{e}{\tau}$
  then $\inferS{C}{\unif\Sv}{e}{\tau}$, $\unif C = C$ and $\unif \tau = \tau$.
\end{theorem}

Note here that the syntax-directed derivation holds with the usage environment $\Sv$ instead of the originally provided environment $\E$. Indeed,
$\E$ does not contain suspended and borrow bindings, since those
are discovered on the fly and recorded in $\Sv$.

The second theorem states that our algorithm is complete: for any given
syntax directed typing derivation, our inference algorithm can find
a derivation that gives a type at least as general.
For this, we need first to provide a few additional definitions.

\begin{definition}
  Given a set of variable $U$ and $\unif$, $\unif'$ and $\phi$
  substitutions. \\
  Then
  $\unif \leq^{\phi}_{U} \unif'$ iff $(\phi \circ \unif)|_{U} = \unif'|_U$.
\end{definition}

\begin{definition}
  Given a constraints $C$ and two schemes
  $\schm = \forall \Multi\tvar. \qual{D}{\tau}$ and
  $\schm' = \forall \Multi\tvar'. \qual{D'}{\tau'} $.
  Then $\entail{C}{\schm \preceq \schm'}$
  iff $\entail{C}{\subst{\tvar}{\tau''} D}$
  and $\entail{C\Cand D'}{\Cleq{\subst{\tvar}{\tau''}\tau}{\tau'}}$
\end{definition}

\begin{definition}
A flattened environment,
noted $\Eflat\E$, is the environment
where all the binders are replace by normal ones. More formally:
\begin{align*}
  \Eflat\E
  &= \left\{ \bvar{x}{\tau} \mid
    \bvar{x}{\tau}\in\E
    \vee \bvar{\borrow{x}}{\borrowty{k}{\tau}}\in\E
    \vee \svar{x}{\tau}^n\in\E
    \right\} \cup \left\{ \bvar{\tvar}{k} \mid \bvar{\tvar}{k}\in\E \right\}
\end{align*}
\end{definition}

We can then state the completeness theorem.

\begin{theorem}[Completeness]
  Let $\inferS{C'}{\unif'\E}{e}{\schm'}$ a typing derivation,
  then $\inferW{\Sv}{(C,\unif)}{\Eflat\E}{e}{\tau}$
  for some substitution $\unif$, constraint $C$ and type $\tau$ such
  that
  \begin{align*}
    (C_\schm,\schm) &= \generalize{C}{\unif\E}{\tau}
    &\unif &\leq^{\phi}_{\fv{\E}} \unif'
    &\entail{C'&}{\phi C}
    &\entail{C'&}{\phi \schm \preceq \schm'}
    % &( C, \schm, \unif) &\leq (C',\schm',\unif')
  \end{align*}
\end{theorem}

The proofs are given in \cref{appendix:infer}.

### 2019-07-08 kindly/sdtyping/extra.tex
State the theorems.

\subsection{Constraint language}

@@

@@

\begin{theorem}
  $\CL$ has the principal constraint property and
  $\operatorname{normalize}$ computes principal normal forms for $\CL$.
\end{theorem}

@@

\begin{lemma}
  $\CL$ is regular, ie, for $x, x'$ two types or kinds,
  $\entail{}{\Ceq{x}{x'}}$ implies
  $\fv{x} = \fv{x'}$
\end{lemma}
\begin{proof}
  The only equalities possibles are between variables (via symmetry) or
  between constants.
\end{proof}

\begin{theorem}
  $\CL$ has the principal constraint property and
  $\operatorname{normalize}$ computes principal normal forms for $\CL$.
\end{theorem}

This is sufficient to show that $HM(\CL)$ is principal. However,
we do not use HM(X) directly, but an extended version with kind inference,
linear and affine types, and borrow.

>>>>

\subsection{Constraint language}
\label{appendix:constraints}

@@

In order to prove this, we first prove that constraint solving
does compute normal forms, and that such normal forms are unique.

@@

We can now prove all the necessary high level properties.

@@

\begin{property}
  $\CL$ is regular, ie, for $x, x'$ two types or kinds,
  $\entail{}{\Ceq{x}{x'}}$ implies
  $\fv{x} = \fv{x'}$
\end{property}
\begin{proof}
  The only equalities possibles are between variables (via symmetry) or
  between constants.
\end{proof}

Finally, we can conclude with all the properties we need for
HM(X):

\begin{theorem}
  $\CL$ has the principal constraint property,
  $\operatorname{normalize}$ computes principal normal forms for $\CL$
  and $\CL$ is regular.
\end{theorem}

This is sufficient to show that $HM(\CL)$ is principal. However,
we do not use HM(X) directly but an extended version with kind inference,
linear and affine types, and borrow.
